
---
title: "Visualization and annotation of read signal over genomic ranges with profileplyr"
author: "Doug Barrows and Tom Carroll"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"

output:
  BiocStyle::html_document:
     number_sections: yes
     toc: true
# bibliography: library.bib
vignette: >
  %\VignetteIndexEntry{Visualization and annotation of read signal over genomic ranges with profileplyr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, results="hide"}
knitr::opts_chunk$set(tidy=FALSE, cache=FALSE,
                      dev="png",
                      message=FALSE, error=FALSE, warning=TRUE)
```

# Introduction
# Import signal quantification from deepTools or soGGi
## Starting with output from deepTools ‘computeMatrix’
## Starting with output from soGGi
# The profileplyr object
## Key components of the object
## Subsetting the profileplyr object
## Changing sample names
## Connecting profileplyr functions using the pipe (%>%) operator
# Export/Conversion of profileplyr object for heatmap visualization of ranges
## Export to a deepTools matrix
## Directly generate a customized EnrichedHeatmap with annotated ranges
## Convert to an EnrichedHeatmap matrix
# Summarize signal for ggplot or heatmap visualization
## Matrix output for heatmaps
## Long output for ggplot
## profileplyr object output with summarized matrix
# Annotating of genomic ranges with clusters, genomic regions, and genes
## K-means and hierarchical clustering of the genomic ranges
## Output matrix with cluster information to deepTools
## Generate group-annotated heatmap in R directly with generateEnrichedHeatmap()
## Visualize mean range signal for each cluster with ggplot
## Gene annotation of ranges
### Annotation of ranges with genes and genomic regions using ChIPseeker
### Annotation of ranges with genes using GREAT
# Grouping ranges by range metadata, gene list, or additional GRanges
## Switch output grouping columns within existing range metadata
## Group by user-supplied GRanges
## Group by user-supplied gene list
### Gene set list contains a character data
### Gene set list with data frames for heatmap annotation (using pipe - %>%)
# Combining multiple profileplyr functions for heatmap annotation
# Acknowledgements
# Session info

```{r sessionInfo}
sessionInfo()
```
